Tracing the Hercules stream around the Galaxy

### 2010-05-19 tex/hercules.tex
groups-simulations: 2d section

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\figurenames}{\figurename s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\tvector}[1]{\boldsymbol{\vec{#1}}}
\newcommand{\vx}{\tvector{x}}
\newcommand{\vv}{\tvector{v}}
\newcommand{\Gaia}{\emph{Gaia}}
\newcommand{\apogee}{APOGEE}
\newcommand{\vR}{\ensuremath{v_R}}
\newcommand{\vphi}{\ensuremath{v_{\phi}}}
\newcommand{\vo}{\ensuremath{v_0}}
\newcommand{\Ro}{\ensuremath{R_0}}
\newcommand{\ro}{\Ro}
\newcommand{\Ab}{\ensuremath{A_b}}
\newcommand{\Rb}{\ensuremath{R_b}}
\newcommand{\Omegab}{\ensuremath{\Omega_b}}
\newcommand{\fdehnen}{\ensuremath{f_{\text{Dehnen}}}}
\newcommand{\sigmaR}{\ensuremath{\sigma_R}}
\newcommand{\rE}{\ensuremath{R_e}}
\newcommand{\Lc}{\ensuremath{L_c}}
\newcommand{\Rs}{\ensuremath{R_s}}
\newcommand{\Rsigma}{\ensuremath{R_{\sigma}}}
\newcommand{\Rolr}{\ensuremath{R_{\text{OLR}}}}

@@

\section{Methodology}

@@

\section{Two-dimensional velocity distributions}

>>>>

\documentclass[12pt,preprint]{aastex}
\usepackage{amssymb,amsmath}
%\usepackage{color,hyperref}
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\figurenames}{\figurename s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\tvector}[1]{\boldsymbol{\vec{#1}}}
\newcommand{\vx}{\tvector{x}}
\newcommand{\vv}{\tvector{v}}
\newcommand{\Gaia}{\emph{Gaia}}
\newcommand{\apogee}{APOGEE}
\newcommand{\hipparcos}{\emph{Hipparcos}}
\newcommand{\vR}{\ensuremath{v_R}}
\newcommand{\vphi}{\ensuremath{v_{\phi}}}
\newcommand{\vo}{\ensuremath{v_0}}
\newcommand{\Ro}{\ensuremath{R_0}}
\newcommand{\ro}{\Ro}
\newcommand{\Ab}{\ensuremath{A_b}}
\newcommand{\Rb}{\ensuremath{R_b}}
\newcommand{\Omegab}{\ensuremath{\Omega_b}}
\newcommand{\fdehnen}{\ensuremath{f_{\text{Dehnen}}}}
\newcommand{\sigmaR}{\ensuremath{\sigma_R}}
\newcommand{\rE}{\ensuremath{R_e}}
\newcommand{\Lc}{\ensuremath{L_c}}
\newcommand{\Rs}{\ensuremath{R_s}}
\newcommand{\Rsigma}{\ensuremath{R_{\sigma}}}
\newcommand{\Rolr}{\ensuremath{R_{\text{OLR}}}}

@@

\section{Methodology}\label{sec:method}

@@

\section{The full planar velocity distribution in the disk}\label{sec:2d}

\citet{dehnen00a} only simulated the velocity distribution at the
position of the Sun for various combinations of the model parameters
defined in the previous Section. Nevertheless, his results on how the
velocity distribution at the Sun changes with the angle between the
Sun-Galactic center line and the bar could also be interpreted as
showing the velocity distribution on the Solar circle for different
Galactocentric azimuths, fixing the bar angle. In this way, these
simulations become predictions for what one might find when one
studies the velocity distribution in locations beyond those probed by
the \hipparcos\ mission.

In \figurename~\ref{fig:rphi2d}, I present the predictions for the
velocity distributions near the Solar circle for the full
360$^{\circ}$ range in Galactocentric azimuths for the fiducial model
from \sectionname~\ref{sec:method}. Only the range $-90^{\circ} \leq
\phi \leq 90^{\circ}$ is shown here, since the barred Galaxy is
symmetric with respect to rotations through 180$^{\circ}$, such that
the predictions for the other side of the Galaxy can be obtained from
those shown by shifting the azimuths by 180$^{\circ}$. The $R = \Ro$
locus is not shown here, but it can be inferred from \figurename~2 in
\citet{dehnen00a}.

It is clear from this \figurename\ that we can trace the Hercules
stream going around the Galaxy and when looking further in or further
out. The effect of the bar is also not confined to a narrow range in
radii, rather the effects of the bar and especially its outer Lindblad
resonance are felt through a large range of radii (0.6 $\leq R/\Ro \leq
1.4$). These effects are strong and robust predictions of the
bar-origin model for the Hercules stream, especially in regions beyond
the Solar radius, where the Hercules feature can be found
significantly removed from the bulk of the velocity distribution,
which might be severely shaped at all locations by the processes that
create the pattern of low-velocity moving groups in the local velocity
distribution (the main mode in \figurename~\ref{fig:obs}). As I
discuss below, \Gaia\ will be able to easily verify these predictions.
